\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

This report drafts the final structure for the project \textit{Parallelization Methods for CNN-Based Computer Vision Inference}. The repository \code{yolo-mpi-people-count} contains two complementary directions. Method 1 is an implemented YOLO/OpenMPI video people-counting pipeline that parallelizes inference at task level over video frames and image tiles. Method 2 is a VGG11 convolution benchmark that is designed to demonstrate feature-map data parallelism with 2D block decomposition and halo exchange.

For Method 1, the report uses measured values already recorded in the project reports. The YOLO pipeline runs on a three-MacBook physical cluster with C++17 and OpenMPI. A video is decomposed temporally into frames and spatially into tiles; each frame-tile pair becomes an MPI task. The final reported pipeline emphasizes static block-cyclic scheduling, rank-local YOLO workers, coordinate remapping, duplicate removal, global NMS, CSV output, and per-rank timing. Dynamic master-worker scheduling is kept only as a comparison point because this workload shows that it increases communication and performs worse.

The recorded YOLO experiments verify serial-versus-MPI correctness on 30 frames with zero mismatched frames. The detector accuracy experiment on 300 MOT17-derived frames reports mean absolute count error 7.983 and predicted average count 8.223 versus ground-truth average 16.207. The input-size experiment selects \(N=600\) frames because the 12-process runtime is 123.667 seconds, satisfying the required two-to-three-minute workload. On \(2N=1200\) frames, the 12-process wall-clock speedup is 1.939x in the original sweep, while two-repeat follow-up measurements report mean 12-process runtime 178.174 seconds and mean speedup 1.781x. Static scheduling outperforms dynamic scheduling on the measured 600-frame 5x4 workload, and weighted static placement improves the idle-gap indicator from 0.466 to 0.235.

For Method 2, the newly published artifacts include VGG11 correctness, input-size, speedup, communication-breakdown, and topology CSV files for blocking halo exchange. The distributed VGG11 path matches the serial classifier result with zero logit error. It scales well from one to four ranks on the \(224 \times 224\) input, reaching 3.569x speedup at four ranks, but becomes slower at eight and twelve ranks because gather time and inter-machine halo communication dominate the small per-rank blocks. A non-blocking halo comparison should be added only after a corresponding CSV artifact is produced.
