\chapter{Problem Description}

\section{CNN-Based Computer Vision Inference}
The shared problem is to execute CNN-based inference faster by distributing work across MPI processes. Let the input be either a video sequence or a CNN feature tensor. In both cases, the program applies a fixed pretrained model; it does not train or update model weights. The parallel system must preserve the output of the serial inference pipeline while reducing runtime or revealing the cost of communication.

\section{Method 1: YOLO Video People Counting}
For the YOLO method, the input video is a finite sequence of frames:
\begin{equation}
    V = \{F_1, F_2, \ldots, F_N\}.
\end{equation}
The target output for each frame is the number of detected people after post-processing:
\begin{equation}
    count(F_i) = |\{b \mid b \text{ is a final person bounding box in } F_i\}|.
\end{equation}

The detector backend is a pretrained YOLO model through a local Python worker process \cite{ultralytics2026}. The MPI program does not split YOLO neural-network kernels. Instead, MPI distributes repeated YOLO calls over frame-tile tasks.

For a tile grid with \(R\) rows and \(C\) columns, frame \(F_i\) is decomposed into:
\begin{equation}
    F_i \rightarrow \{T_{i,1}, T_{i,2}, \ldots, T_{i,R \times C}\}.
\end{equation}
A task is one frame-tile inference:
\begin{equation}
    task(i,j) = T_{i,j}, \qquad |\mathcal{T}| = N \times R \times C.
\end{equation}
Rank 0 converts tile-local detections back to full-frame coordinates, applies tile ownership filtering, removes duplicates, and writes final frame counts.

\section{Method 2: VGG11 Convolution Benchmark}
For the VGG11 method, the input is a CNN feature map. Each convolution layer receives a tensor and computes an output tensor using a fixed \(3 \times 3\) kernel. The repository uses VGG11 without BatchNorm, with scaled channel profiles such as \code{tiny}, \code{small}, and \code{full}. The purpose is to demonstrate data parallelism inside CNN convolution layers.

For each convolution layer, \(P\) MPI ranks are arranged into a \(P_r \times P_c\) process grid. The spatial dimensions \(H \times W\) of the feature map are split into 2D blocks. Each rank computes the convolution output for its local core block. Since a \(3 \times 3\) convolution needs neighboring pixels, ranks must exchange one-pixel halo rows, columns, and corners before local computation.

Method 1 and Method 2 therefore exercise different parallel-programming patterns:
\begin{table}[H]
\centering
\caption{Comparison of the two project methods}
\label{tab:method-comparison}
\begin{tabularx}{\textwidth}{L{3.2cm} X X}
\toprule
Aspect & Method 1: YOLO video inference & Method 2: VGG11 convolution \\
\midrule
Parallelism level & Task parallelism & Data parallelism \\
Decomposition & Temporal-spatial frame/tile tasks & 2D feature-map blocks \\
Mapping & 1D flattened task mapping & 2D block/process-grid mapping \\
Communication & Master-worker dispatch and result collection & Neighbor halo exchange \\
Topology & Star topology around rank 0 & 2D Cartesian/mesh-like topology \\
Correctness target & Serial count equals MPI count & Serial convolution/VGG stack equals distributed output \\
\bottomrule
\end{tabularx}
\end{table}

\section{Correctness Targets}
The serial YOLO baseline processes the same frame-tile task list without MPI distribution and applies the same post-processing pipeline. Parallel correctness is:
\begin{equation}
    \Delta_i = |count_{parallel}(F_i) - count_{serial}(F_i)|.
\end{equation}
The VGG11 correctness target is numerical agreement between the serial convolution stack and the distributed convolution stack, typically reported as maximum absolute error or pass/fail under a chosen tolerance.
