\chapter{Parallelization Method}

\section{Level of Parallelism}
The implementation exploits task-level parallelism over independent frame-tile inference tasks. It also uses data decomposition because the input video is split temporally into frames and spatially into tiles. The neural-network inference inside YOLO11n is not parallelized by this project; it is treated as local work performed by each active MPI rank.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    node distance=0.9cm and 1.1cm,
    box/.style={draw, rounded corners=2pt, align=center, minimum height=0.8cm, minimum width=2.25cm, fill=gray!8},
    rank/.style={draw, rounded corners=2pt, align=center, minimum height=0.75cm, minimum width=1.95cm, fill=blue!7},
    arrow/.style={-{Latex[length=2mm]}, thick}
]
\node[box] (video) {Input video};
\node[box, right=of video] (tasks) {Frame-tile\\task list};
\node[rank, right=of tasks, yshift=1.0cm] (r0) {Rank 0\\scheduler};
\node[rank, right=of tasks] (r1) {Rank 1\\YOLO worker};
\node[rank, right=of tasks, yshift=-1.0cm] (r2) {Rank 2..P-1\\YOLO workers};
\node[box, right=1.2cm of r1] (agg) {Rank 0\\NMS + CSV};
\draw[arrow] (video) -- (tasks);
\draw[arrow] (tasks) -- (r0);
\draw[arrow] (tasks) -- (r1);
\draw[arrow] (tasks) -- (r2);
\draw[arrow] (r0) -- (agg);
\draw[arrow] (r1) -- (agg);
\draw[arrow] (r2) -- (agg);
\end{tikzpicture}
\caption{System architecture for offline YOLO-MPI people counting.}
\label{fig:architecture}
\end{figure}

\section{Temporal-Spatial Data Decomposition}
The input is first decomposed across time and then across image space. For \(N\) frames and an \(R \times C\) tile grid, the task set contains:
\begin{equation}
    |\mathcal{T}| = N \times R \times C.
\end{equation}
The average task granularity per rank is:
\begin{equation}
    G = \frac{N \times R \times C}{P},
\end{equation}
where \(P\) is the number of MPI ranks. Larger tile grids create more tasks and may improve scheduling flexibility, but they can increase duplicated border work and post-processing overhead.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    tile/.style={draw, minimum width=1.1cm, minimum height=0.55cm, fill=green!7},
    arrow/.style={-{Latex[length=2mm]}, thick}
]
\node at (0,1.2) {\(V=\{F_1,\ldots,F_N\}\)};
\foreach \x/\lab in {0/\(F_1\),1.3/\(F_2\),2.6/\(\cdots\),3.9/\(F_N\)} {
  \node[draw, minimum width=1.0cm, minimum height=0.7cm, fill=gray!8] at (\x,0.45) {\lab};
}
\draw[arrow] (1.95,0.0) -- (1.95,-0.65);
\node at (1.95,-0.95) {\(F_i \rightarrow \{T_{i,1},\ldots,T_{i,R C}\}\)};
\foreach \r in {0,1,2} {
  \foreach \c in {0,1,2,3} {
    \node[tile] at (-0.15+\c*1.1,-1.55-\r*0.55) {};
  }
}
\node at (5.5,-2.1) {each tile \(T_{i,j}\) becomes one MPI task};
\draw[arrow] (3.35,-2.1) -- (4.35,-2.1);
\end{tikzpicture}
\caption{Temporal-spatial task decomposition from frames to frame-tile tasks.}
\label{fig:task-decomposition}
\end{figure}

\section{1D Task Mapping and Static Scheduling}
Static scheduling flattens all tasks into a one-dimensional list. For task index \(i\), chunk size \(k\), and \(P\) ranks, the owner is:
\begin{equation}
    owner(task_i) = \left\lfloor \frac{i}{k} \right\rfloor \bmod P.
\end{equation}
With \(k=1\), this becomes round-robin assignment. Static scheduling has low coordination cost because each rank can decide its own tasks locally, but it can suffer when expensive tiles are unevenly distributed.

\section{Dynamic Master-Worker Scheduling}
Dynamic scheduling uses rank 0 as a master. Rank 0 owns the remaining task queue, dispatches work to available ranks, receives completed result payloads, and sends the next task. If \code{master\_compute} is enabled, rank 0 may also process local tasks while waiting for worker results.

\section{Communication Strategy and Star Topology}
The dynamic scheduler uses a star topology: workers communicate with rank 0, not directly with each other. In static scheduling, ranks work independently and communicate mainly during final aggregation.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    master/.style={draw, circle, minimum size=1.4cm, fill=blue!10, align=center},
    worker/.style={draw, rounded corners=2pt, minimum width=1.6cm, minimum height=0.7cm, fill=gray!8, align=center},
    arrow/.style={<->, >=Latex, thick}
]
\node[master] (m) at (0,0) {Rank 0\\master};
\node[worker] (w1) at (90:3.0cm) {Rank 1};
\node[worker] (w2) at (30:3.0cm) {Rank 2};
\node[worker] (w3) at (-30:3.0cm) {Rank 3};
\node[worker] (w4) at (-90:3.0cm) {Rank 4};
\node[worker] (w5) at (210:3.0cm) {Rank P-1};
\node[worker] (w6) at (150:3.0cm) {Other ranks};
\foreach \name in {w1,w2,w3,w4,w5,w6} {
  \draw[arrow] (m) -- (\name);
}
\end{tikzpicture}
\caption{Master-worker star communication topology.}
\label{fig:communication-topology}
\end{figure}

\section{Blocking Communication and MPI Polling}
The implementation uses blocking message transfer for task and result payloads. Task dispatch uses \code{MPI\_Send}; task reception and result reception use \code{MPI\_Recv}; static aggregation uses \code{MPI\_Gather} and \code{MPI\_Gatherv}. The dynamic scheduler also calls \code{MPI\_Iprobe} so rank 0 can poll whether a worker result is ready before choosing whether to compute locally. This is polling only: the actual payload transfer is still blocking, and no \code{MPI\_Isend} or \code{MPI\_Irecv} payload path is reported.

\section{Load Balancing Considerations}
Per-rank time is interpreted as compute time, communication time, and idle/wait time:
\begin{equation}
    T_r = T_{compute,r} + T_{comm,r} + T_{idle,r}.
\end{equation}
The imbalance metric used in the report is:
\begin{equation}
    imbalance = \frac{T_{max} - T_{min}}{T_{max}}.
\end{equation}
Dynamic scheduling should help when task costs vary, while static scheduling should have lower scheduling overhead. The final conclusion must be based on measured CSV files rather than expectation.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    task/.style={draw, minimum width=0.65cm, minimum height=0.45cm, fill=green!10},
    rank/.style={draw, rounded corners=2pt, minimum width=1.0cm, minimum height=0.45cm, fill=blue!7},
    arrow/.style={-{Latex[length=2mm]}, thick}
]
\node at (0,1.1) {Static};
\foreach \i in {0,...,7} {\node[task] at (-2.8+\i*0.7,0.45) {\(\i\)};}
\foreach \r/\y in {0/-0.2,1/-0.75,2/-1.3} {
  \node[rank] at (-3.7,\y) {R\r};
}
\draw[arrow] (-2.8,0.2) -- (-3.0,-0.2);
\draw[arrow] (-2.1,0.2) -- (-3.0,-0.75);
\draw[arrow] (-1.4,0.2) -- (-3.0,-1.3);

\node at (3.2,1.1) {Dynamic};
\node[draw, rounded corners=2pt, fill=gray!8, minimum width=2.3cm] (q) at (3.2,0.45) {Rank 0 queue};
\foreach \r/\y in {1/-0.2,2/-0.75,3/-1.3} {
  \node[rank] (d\r) at (1.9,\y) {R\r};
  \draw[arrow] (q) -- (d\r);
}
\node[align=left] at (5.25,-0.75) {next task goes\\to free rank};
\end{tikzpicture}
\caption{Static assignment versus dynamic task dispatch.}
\label{fig:static-dynamic}
\end{figure}

\section{Post-processing and NMS Deduplication}
After inference, detections are converted from tile coordinates to frame coordinates. Rank 0 applies tile ownership filtering, confidence sorting, global non-maximum suppression, and duplicate rules for boxes near tile boundaries. The final frame count is the number of remaining person boxes.

\section{Parallel Algorithm Pseudocode}
\begin{algorithm}[H]
\caption{Serial YOLO people-counting baseline}
\begin{algorithmic}[1]
\STATE Build task list \(\mathcal{T}\) from \(N\) frames and \(R \times C\) tiles.
\FOR{each task \(T_{i,j}\in\mathcal{T}\)}
    \STATE Run local YOLO inference on tile \(T_{i,j}\).
    \STATE Convert detections to frame coordinates.
\ENDFOR
\STATE Apply tile ownership filtering and global NMS per frame.
\STATE Write \code{frame\_counts.csv}, \code{bboxes.csv}, and \code{summary.csv}.
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{Static MPI YOLO inference}
\begin{algorithmic}[1]
\STATE All ranks build the same task list \(\mathcal{T}\).
\FOR{each task index \(i\)}
    \IF{\(owner(task_i)=\lfloor i/k \rfloor \bmod P\) equals local rank}
        \STATE Run YOLO inference and store local detections and timing.
    \ENDIF
\ENDFOR
\STATE Gather serialized detections and rank metrics to rank 0.
\STATE Rank 0 applies post-processing and writes CSV outputs.
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{Dynamic master-worker MPI YOLO inference}
\begin{algorithmic}[1]
\IF{rank is 0}
    \STATE Initialize task queue and dispatch initial tasks to workers.
    \WHILE{unfinished tasks or active workers remain}
        \STATE Poll for ready worker results using \code{MPI\_Iprobe}.
        \STATE Receive completed payloads with blocking \code{MPI\_Recv}.
        \STATE Send next task or termination message with \code{MPI\_Send}.
        \STATE Optionally compute one local task if \code{master\_compute} is enabled.
    \ENDWHILE
    \STATE Post-process all detections and write CSV outputs.
\ELSE
    \WHILE{a task is received}
        \STATE Run YOLO inference and send result payload to rank 0.
    \ENDWHILE
\ENDIF
\end{algorithmic}
\end{algorithm}
