\chapter{Conclusion}

This project implemented and evaluated a task-parallel people-counting pipeline using YOLO, C++17, and OpenMPI on a three-machine physical cluster. The input video is decomposed both temporally and spatially: frames create independent work over time, while image tiles create finer-grained work inside each frame. These frame-tile tasks are assigned through rank-based mapping and weighted-node mapping, processed locally by YOLO workers, and merged through coordinate remapping, tile-owner filtering, and global duplicate removal.

The project demonstrates the main parallel-programming concepts required by the course. The level of parallelism is task parallelism. The decomposition technique is hybrid temporal-spatial decomposition. The mapping strategy includes a simple rank-based mapping and a weighted-node mapping designed for a heterogeneous cluster. The communication pattern is either a centralized gather to rank 0 or a hierarchical node-local merge followed by a compact global gather. Load balance is measured with an idle-gap indicator, and the best measured placement is the weighted $4/6/2$ configuration.

The experiments show that the parallel implementation preserves the behavior of the serial tiled pipeline. In the correctness experiment, the MPI run has zero mismatched frames, zero maximum count error, and zero mean count error compared with the serial pipeline. This verifies parallel correctness, although detector accuracy against human ground truth remains a separate limitation of the pretrained YOLO model.

For input-size selection, the selected course workload is

\[
N = 600 \text{ frames}
\]

with a $5 \times 4$ tile grid and 12 MPI processes. This produces 12,000 frame-tile inference tasks and a wall-clock runtime of 123.667 seconds, satisfying the required 120--180 second interval. For the strong-scaling experiment, the workload is doubled to

\[
2N = 1200 \text{ frames}.
\]

The final measured wall-clock speedup is 1.939x at $P = 12$, while the compute-oriented curve stays slightly better at about 2.011x. The result is useful but sublinear, mainly because of communication overhead, rank-0 aggregation, duplicate removal, tile-boundary effects, and heterogeneous node speeds.

The granularity and mapping experiments show that finer tiling creates more parallel work units but also increases communication and post-processing cost. Therefore, more tasks do not automatically produce better performance. The weighted $4/6/2$ placement improves load balance by reducing the idle-gap indicator from 0.466 under uniform $4/4/4$ placement to 0.235, which is below the 25\% imbalance threshold.

The project also exposes practical limitations. Centralized merging at the master can become a bottleneck, tiled inference can reduce visual context near region boundaries, and the pretrained detector may undercount crowded scenes. Even with these limitations, the system provides a clear demonstration of distributed computer-vision inference with MPI, including decomposition, mapping, communication, load balancing, correctness verification, granularity analysis, and speedup measurement.
